Errorless irrationality: removing error-driven components from the inverse base-rate effect paradigm

### 2023-01-27 docs/manuscript.tex
transitional commit before final submission

There are alternative theories of the IBRE that do not rely on processes that calculate prediction error.
A version of the dissimilarity-similarity generalized context model \citeA<DGCM, >{stewart2007dissimilarity} modified by \citeA{o2018model} has been proposed as an explanation of the IBRE.
From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strength of exemplars and dissimilarity from exemplars (stored category representations).
$BC \to rare$ arises due to the combination of two factors: the high salience of $C$ relative to $B$ and the impact of the dissimilarity of $BC$ to the common category exemplar on the decision process.
Nonetheless, the model can accommodate these results only informally because it does not specify the mechanisms which encode information and produce the attentional values of each stimulus.
DGCM is a model of the test phase.
In that sense, the current experimental manipulations cannot be represented in the model.
This is the same challenge we encountered with process models - the specifications of these theories are unable to incorporate the changes to the experimental procedure.

@@

Across two experiments, we investigated whether the demonstration of the IBRE requires the "guess-and-correct" components of the standard experimental procedure.
In Experiment 1, we conducted a successful conceptual replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.
In addition, Experiment 2 further suggests that the IBRE generalizes beyond simple predictive-learning \cite<e.g. >{medin1988problem,kruschke1996base,wills2014attention} and decision-making \cite{johansen2007paradoxical} paradigms.
This further suggests prediction error in terms of explicit feedback is not a necessary condition.
Theories of IBRE are inadequate to account for these findings, largely because of their inability to extend beyond supervised learning.

>>>>

There are alternative theories of the IBRE that do not rely on processes that calculate prediction error.
A version of the dissimilarity-similarity generalized context model \citeA<DGCM, >{stewart2007dissimilarity} modified by \citeA{o2018model} has been proposed as an explanation of the IBRE.
From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strength of exemplars and dissimilarity from exemplars (stored category representations).
$BC \to rare$ arises due to the combination of the following factors: the high salience of $C$ relative to $B$ and the impact of the dissimilarity of $BC$ to the most remembered common category exemplar on the decision process.
Nonetheless, the model can accommodate these results only informally because it does not specify the mechanisms which encode information and produce the attentional values of each stimulus.
DGCM is a model of the test phase.
In that sense, the current experimental manipulations cannot be represented in the model.
This is the same challenge we encountered with process models - the specifications of these theories are unable to incorporate the changes to the experimental procedure.

@@

Across two experiments, we investigated whether the demonstration of the IBRE requires the prediction-and-feedback components of the standard experimental procedure.
In Experiment 1, we conducted a successful conceptual replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.
In addition, Experiment 2 further suggests that the IBRE generalizes beyond simple predictive-learning \cite<e.g. >{medin1988problem,kruschke1996base,wills2014attention} and decision-making \cite{johansen2007paradoxical} paradigms.
This further suggests prediction error in terms of explicit feedback is not a necessary condition.
Theories of IBRE are inadequate to account for these findings, largely because of their inability to extend beyond supervised learning.
